---
title: "Making data for OLS of AEDARs"
---

<!-- I recommend modifying/looking at this file in RStudio -->

<!-- When this file "compiles" (using the knit button in RStudio, it reads in the _output.yaml file in the same directory as this file. Alternatively, you could put the metadata from the .yaml file in this file between the ---) -->

```{r opt-chunks, echo=F}
knitr::opts_chunk$set(cache=2,
                      cache.rebuild = T, 
                      message = F, warning = F,
                      autodep=TRUE,cache.lazy = FALSE,comment = "")
```


```{r libraries}
library(dplyr)
```

This notebook processes the "full" AEDAR dataset and subsets it into a smaller, more manageable dataset for regressing age of reports. Only valproate-containing AEDARs will be used as a proof of concept.

1. Upload table amd filter reports for thosereporting a Valproate ADR.

```{r,  table_upload}

stats <- data.table::fread("zcat < ../data/aeolous_stats_a_gte_20.tsv.gz",sep="\t")

valproate <- stats %>% filter(drug_concept_name=="Valproate")

```

2. Spread out AEDARs as columns, reports as rows and frequency as values

```{r aedar_spread}

joined <- valproate %>% 
  count(id,drug_outcome_name) %>% 
  tidyr::spread(drug_outcome_name,n) %>% 
  left_join(valproate,by = "id") 

df_master_age <- joined %>% 
  select(id:Valproate_Wound,master_age)

```

3. Write matrix

```{r write}

readr::write_csv(df_master_age,"../data/valproate_adr_sparse_matrix_master_age.csv")

system("gzip ../data/valproate_adr_sparse_matrix_master_age.csv")
```

Viola. Now we can ask if valproate-containing AEDARs are dependent on age for all patients (based on reports) that took valproate. We can use OLS in python since that may work better than in R, but I would have to test it out. 
